# Results

```{r}
library(tidyverse)
dota2 <- read.csv("addata.csv")
```

## Select Ability Phase
Before answering the research questions, we need to select the abilities we would like to analyze. There are over 500 abilities in total, which is an extremely large pool, so we decide to choose 30 of them with the most sufficient data given. Thus, firstly, we will draw the barplot showing the frequencies of all abilities, and our goal is to select the top 30 to do the future analysis. 

```{r}
df = as.data.frame(table(dota2$ability))
names(df)[1] = 'ability'
df$rank = rank(df$Freq)
df <- df[order(df$rank,decreasing=TRUE),]
df <- head(df,30)
```

```{r}
library(ggplot2)
library(forcats)
ggplot(df, aes(x=reorder(ability,-Freq),y=Freq))+
  geom_col(color='blue', fill ='cornflowerblue', width=0.8)+
  labs(title="Frequency Count of Top 30 Abilities",x="Ability", y = "Count")+
  theme(axis.text.x = element_text(angle = 90, size = 8)) 

dota2 <- dota2[dota2$ability %in% df$ability,]
```

The abilities are ordered based on their frequencies in the dataset, and the above graph only shows the abilities with top 30 frequencies. We can see for these top abilities, there are not much variation in their frequencies, all around 30 counts. Thus, we can ensure that there's sufficient data to analyze these abilities.

## First Research Question

- Which abilities have the highest damage, kills, gold, experiences points earned, and kda (a statistic that is used to measure a player's performance), and are therefore the most valuable?

To answer the question, we decide to draw parallel coordinate plots which is useful when we want to get the abilities that top the charts in all five indicators. 

```{r}
library(GGally)
library(dplyr)
colnames(dota2)[35] <- "damage"
df_ability <- dota2[,c("total_gold","kills","kda","damage", "total_xp", "ability")]
GGally::ggparcoord(data = df_ability, columns = 1:5, alphaLines = .4,
                   groupColumn = 6, splineFactor = TRUE, title = "Parallel Coordinates Plot")
```
This is the parallel coordinate plot colored by abilities so that we can see the trend of lines of different colors. In general, the lines are compact at the start and at the end, while it's relatively widespread in the middle, with several lines having extremely large value of kda (potential outliers) that belong to the pink category "skywrath_mage_concussive_shot". 

```{r}
library(parcoords)
df <- dota2[,c("total_gold","kills","kda","damage", "total_xp")]
parcoords::parcoords(data = df, rownames = FALSE, brushMode = "1d-axes")
```

This is the interactive parallel coordinate plot for all 30 abilities. We make it interactive such that we can move axis around and we can filter for particular section which easier to see the top abilities. However, since the dataset is too large, the lines overlap, so we need to select specific region to interpret the plot. 

```{r}
df <- df[df$kills > 16 & df$total_xp > 36000  & df$total_gold > 32000 & df$kda > 10,]
parcoords::parcoords(data = df, rownames = FALSE, brushMode = "1d-axes")
```
From the interactive parallel coordinate plot, I select a specific region where these abilities have relatively high values across all five variables, damage, kills, gold, experiences points earned, and kda, and I filter the dataset based on that. However, it's important to notice that in the above plot there are only three lines because two abilities have exactly same all five attributes so the two lines overlap. 

```{r}
df_ability <- df_ability[df_ability$kills > 16 & df_ability$total_xp > 36000  & df_ability$total_gold > 32000 & df_ability$kda > 10 ,]
df_ability
```

In total, I select 4 top abilities through these parallel coordinate plots, which are the following:\
1. omniknight_hammer_of_purity\
2. sandking_caustic_finale\
3. drow_ranger_marksmanship\
4. rattletrap_battery_assault\

## Second Research Question

- Do these four good abilities function better on radiant or dire?

To answer the question, we decide to draw mosaic plot here because it can incorporate group size into account and can clearly show the relationship between multiple categorical variables, which are win (1 is win and 0 is lose), isRadient(True is radient and False is dire), and ability in our case.

```{r}
library(vcd)
library(RColorBrewer)
df_2 <- dota2[dota2$ability == "omniknight_hammer_of_purity" | dota2$ability == "sandking_caustic_finale" | dota2$ability == "drow_ranger_marksmanship" | dota2$ability == "rattletrap_battery_assault",]
df_2[df_2 == "omniknight_hammer_of_purity"] <- 'o'
df_2[df_2 == "sandking_caustic_finale"] <- 's'
df_2[df_2 == "drow_ranger_marksmanship"] <- 'd'
df_2[df_2 == "rattletrap_battery_assault"] <- 'r'

vcd::mosaic(win~isRadiant+ability, data = df_2, direction = c("v","v","h"), highlighting_fill = brewer.pal(4, 'Set3'),) 
```

In this mosaic plot, label 'd' represents the ability "drow_ranger_marksmanship"; 'r' represents the ability "rattletrap_battery_assault"; 'o' represents the ability "omniknight_hammer_of_purity"; 's' represents the ability "sandking_caustic_finale". These abbreviations make the plot clearer to read by avoiding label overlaps.\
From the above mosaic plot, we can see for the three abilities "omniknight_hammer_of_purity", "drow_ranger_marksmanship", and "rattletrap_battery_assault", they perform significantly better on radiant, while for the ability "sandking_caustic_finale", it performs better on dire (not radiant), so the choice of radiant/dire really depends on the abilities chosen. In our case, these four abilities generally function better on radiant. 

## Third Research Question

- What heroes are most commonly associated with the four good abilities?

This question can be answered by drawing the separate barplots for each ability. When people are assigned hero, these heroes are designed with a specific role in mind, so they have tendence to select correlated abilities. From the barplots we can know what heroes are highly correlated to the selection of each ability among all four. 

```{r}
df_d <- df_2[df_2$ability == "d",] 
df_d = as.data.frame(table(df_d$hero))
names(df_d)[1] = 'hero'
df_d$rank = rank(df_d$Freq)
df_d <- df_d[order(df_d$rank,decreasing=TRUE),]

a = ggplot(df_d, aes(x=reorder(hero,-Freq),y=Freq))+
  geom_col(color='pink', fill ='lightpink', width=0.8)+
  labs(title="Count of Hero Associated with 'd'",x="hero", y = "Count")+
  theme(axis.text.x = element_text(angle = 90, size = 8)) 

df_r <- df_2[df_2$ability == "r",] 
df_r = as.data.frame(table(df_r$hero))
names(df_r)[1] = 'hero'
df_r$rank = rank(df_r$Freq)
df_r <- df_l[order(df_r$rank,decreasing=TRUE),]

b = ggplot(df_r, aes(x=reorder(hero,-Freq),y=Freq))+
  geom_col(color='pink', fill ='lightpink', width=0.8)+
  labs(title="Count of Hero Associated with 'r'",x="hero", y = "Count")+
  theme(axis.text.x = element_text(angle = 90, size = 8)) 

df_o <- df_2[df_2$ability == "o",] 
df_o = as.data.frame(table(df_o$hero))
names(df_o)[1] = 'hero'
df_o$rank = rank(df_o$Freq)
df_o <- df_o[order(df_o$rank,decreasing=TRUE),]

c = ggplot(df_o, aes(x=reorder(hero,-Freq),y=Freq))+
  geom_col(color='pink', fill ='lightpink', width=0.8)+
  labs(title="Count of Hero Associated with 'o'",x="hero", y = "Count")+
  theme(axis.text.x = element_text(angle = 90, size = 8)) 

df_s <- df_2[df_2$ability == "s",] 
df_s = as.data.frame(table(df_s$hero))
names(df_s)[1] = 'hero'
df_s$rank = rank(df_s$Freq)
df_s <- df_s[order(df_s$rank,decreasing=TRUE),]

d = ggplot(df_s, aes(x=reorder(hero,-Freq),y=Freq))+
  geom_col(color='pink', fill ='lightpink', width=0.8)+
  labs(title="Count of Hero Associated with 's'",x="hero", y = "Count")+
  theme(axis.text.x = element_text(angle = 90, size = 8)) 
```

```{r}
library("ggpubr")
ggarrange(a,b,c,d,common.legend = TRUE, legend = "bottom")
```
From above barplots, we can see the hero Drow Ranger tends to be highly associated with the ability "drow_ranger_marksmanship"; the hero Lina is highest correlated to the ability "rattletrap_battery_assault"; the hero Omniknight is highest correlated to the ability "omniknight_hammer_of_purity"; the hero Gyrocoopter is highest correlated to the ability "sandking_caustic_finale". When people are randomly assigned these heroes, they are more likely to also select these four good abilities with a significant frequency difference compared to other heroes. 


## Fourth Research Question

- How many resources are consumed by these four abilities compared to the the other abilities?\

We want to draw histograms of continuous variable net worth faceted by all 30 abilities to see the distribution of net resources of each player possessing good abilities compared to rest of the abilities. 

```{r}
dota2[dota2 == "omniknight_hammer_of_purity"] <- 'o'
dota2[dota2 == "sandking_caustic_finale"] <- 's'
dota2[dota2 == "drow_ranger_marksmanship"] <- 'd'
dota2[dota2 == "rattletrap_battery_assault"] <- 'r'
ggplot(dota2, aes(net_worth))+
  geom_histogram(color = 'cornflowerblue', fill = 'lightblue', bins=15)+
  facet_wrap(~ability)+
  labs("Faceted Histogram of Net Worth")+
  theme_grey(9)
```
From these histograms, we can see the three abilities "omniknight_hammer_of_purity", "drow_ranger_marksmanship", and "rattletrap_battery_assault" consume relative moderate amount of resources compared to the rest of the abilities, while the fourth ability "sandking_caustic_finale" spend high amount of resources with a widespread value range on net worth.
